---
title: "Elevation Data Processing"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = '/Users/Hao/Dropbox/Private/Research/DIFM/processDEM/')
```

## Load required packages
```{r,message=FALSE, warning=FALSE}
library(raster)
library(rgdal)
library(gstat)
```

## Preprocessing elevation data
### Read from shapefile and define geographc coordinate system as WGS84
```{r,message=FALSE, warning=FALSE}
bohnhoff <- readOGR("data", "bohnhoff-adams-2016-ec")
proj4string(bohnhoff) = CRS("+init=epsg:4326")
```

### Reproject the data to WGS84/UTM zone 15N
```{r}
bohnhoff.prj <- spTransform(bohnhoff, CRS("+proj=utm +zone=15 +datum=WGS84"))
```

### Create Digital Elevation Model (DEM) raster from points
#### Method without interpolation
```{r}
g <- raster(bohnhoff.prj)
res(g) <- c(15, 15) # set the resolution to 15 meters
bohnhoff.dem <- rasterize(bohnhoff.prj, g, field = bohnhoff.prj$ELEVATION, fun = mean, na.rm = TRUE)
```
#### Method with interpolation (IDW - Inverse Distance Weight)
```{r}
g <- raster(bohnhoff.prj)
res(g) <- c(5, 5) # the resolution can be finer
bohnhoff.dem <- rasterize(bohnhoff.prj, g, field = bohnhoff.prj$ELEVATION, fun = mean, na.rm = TRUE)
r.idw <- gstat(id = "ELEVATION", formula = ELEVATION~1, data=bohnhoff.prj, 
      nmax=10, set=list(idp = .5)) # idp is the power of inverse distance function
bohnhoff.dem <- interpolate(bohnhoff.dem, r.idw)
```

### Plot the raster
```{r fig.cap="Elevation Map", fig.width=8, fig.height=6, fig.align='center'} 

plot(bohnhoff.dem)

```

```{r fig.cap="Elevation Map (Overlap with original points)", fig.width=8, fig.height=6, fig.align='center'} 

plot(bohnhoff.dem)
plot(bohnhoff.prj,pch=20,add=T, cex=0.5)

```


## Generate derived products from elevation data
+ TPI: Topographic Position Index - the difference between the value of a cell and the mean value of its 8 surrounding cells
+ TRI: Terrain Ruggedness Index - the mean of the absolute differences between the value of a cell and the value of its 8 surrounding cells
+ Roughness:  the minimum value of a cell and its 8 surrounding cells
```{r}
bohnhoff.derived <- terrain(bohnhoff.dem, opt=c('slope', 'aspect','flowdir','TPI','TRI','roughness'), unit='degrees')
```
### Plot derived data
```{r fig.cap="Derived data from DEM", fig.width=10, fig.height=6, fig.align='center'} 

plot(bohnhoff.derived)

```